---
title: "fireSense_SpreadFit"
author: "Jean Marchal (jean.d.marchal@gmail.com)"
date: "`r format(Sys.time(), '%B %Y')`"
output:
  html_document: default
  pdf_document: default
---

# Overview
Fit statistical models that can be used to parameterize the fire spread component of simulation models (e.g. fireSense). This module implement a Pattern Oriented Modelling (POM) approach to derive spread probabilities from final fire sizes. Spread probabilities can vary between pixels, and thus reflect local heterogeneity in environmental conditions.

# Download the module
```{r download module, eval = FALSE, echo = TRUE}
library(SpaDES)

moduleName <- "fireSense_SpreadFit"

spadesModulesDirectory <- tempdir() # Location where the module will be downloaded

downloadModule(moduleName, path = spadesModulesDirectory)
```

# Usage
## Module parameters
Name|Default|Description
----|:-------|---------------------------------------------------------------------
`formula`|`NA`|an object of class formula: a symbolic description of the model to be fitted. Only the RHS needs to be provided.
`data`|`"dataFireSense_SpreadFit"`|a character vector indicating the names of objects in the `simList` environment in which to look for variables present in the model formula. `data` objects can be RasterLayers or RasterStacks (for time series). If variables are not found in `data` objects, they are searched in the `simList` environment.
`fires`|`"fires"`|an object of class SpatialPointsDataFrame describing fires starting locations, final sizes ('size' column), and possibly the starting dates ('date' column) if fires are to be spread at different time intervals. If the 'date' column is not present, all fires are assumed to have started at the same time interval.
`lower`|`NULL`|see `?DEoptim`.
`upper`|`NULL`|see `?DEoptim`.
`itermax`|`500`|integer defining the maximum number of iterations allowed (DEoptim optimizer).
`nCores`|`1`|non-negative integer. Defines the number of logical cores to be used for parallel computation. The default value is 1, which disables parallel computing.
`trace`|`0`|non-negative integer. If > 0, tracing information on the progress of the optimization are printed every `trace` iteration. Default is 0, which turns off tracing.
`initialRunTime`|`start(simList)`|when to start this module? By default, the start time of the simulation.
`intervalRunModule`|`NA`|optional. Interval between two runs of this module, expressed in units of simulation time.
|||

## Usage example
```{r module usage example, eval = FALSE}
library(SpaDES)

# Define simulation parameters
times <- list(start = 1, end = 1, timeunit = "year")
modules <- list("fireSense_SpreadFit")
paths <- list(
  modulePath = spadesModulesDirectory
)

# Define module parameters
parameters <- list(
  fireSense_SpreadFit = list(
    formula = ~ weather + vegetation - 1,
    lower = c(.2, .1, .01, .3, 0.001, 0.001),
    upper = c(.5, 10, .2, 4, .3, .3),
    trace = 5,
    nCores = 8,
    itermax = 5
  )
)

# Define from where and how data will be loaded in the simList environment
inputs <- data.frame(
  objectName = c("fires", "beta", "theta"),
  file = c("fires.shp", "weather.tif", "vegetation.tif"),
  fun = c("shapefile", "stack", "stack"),
  package = c("raster", "raster", "raster"),
  loadTime = 1
)

# Create the simList
sim <- simInit(
  times = times,
  modules = modules,
  paths = paths,
  params = parameters,
  inputs = inputs
)

tweakSimList <- function(sim) {
  loadFiles(sim)
  sim[["fires"]][["date"]] <- NULL
  sim[["beta"]] <- setNames(raster::unstack(sim[["beta"]])[[1]], "beta")
  sim[["theta"]] <- setNames(raster::unstack(sim[["theta"]])[[1]], "theta")
  inputs(sim) <- list()
  sim
}

# Fires spread at different times
sim2 <- spades(sim)
sim2$fireSense_SpreadFitted

# All fires at the same time
sim <- tweakSimList(sim)

sim3 <- spades(sim)
sim3$fireSense_SpreadFitted
```

# Events
Events are scheduled as follows:

- Module initiation
- Model fitting

# Data dependencies
## Input data
- **dataFireSense_SpreadFit**: One or more objects of class 'RasterLayer' or 'RasterStack', in which to look for variables present in the model formula. RasterStacks can be used in cases where fires have started at different times and should not be spread at the same time interval, but are still used to describe the same fire size distribution. In this case, the number of layers in the RasterStack should equal the number of distinct dates in column 'date'.
- **fires**: an object of class SpatialPointsDataFrame describing fires starting locations, final sizes ('size' column), and possibly the starting dates ('date' column) if fires are to be spread at different time intervals. If the 'date' column is not present, all fires are assumed to have started at the same time interval.

## Output data
- **fireSense_SpreadFitted**: an object of class `fireSense_SpreadFit`, i.e. a list containing the following components:

    - formula
    - coef (A, B, D, G are parameters of the logistic function)
    - AD (Two-samples Anderson-Darling statistic)

# Links to other modules
This module can be used in association with fireSense_SpreadPredict to derive fire spread probabilities that are sensistive to environmental conditions.

